### 2019-06-12 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Fixed typos in notations of Algorithms RandTopM/MCTopM in Ch5

% % FIXME
% \TODOL{Est-ce qu'on enlève toute mention de la lower-bound et de "order-optimal" du coup ?!}
% % FIXME

@@

\TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T - 1$}{
                  %
                  \eIf{$A^j(t) \notin \TopM(t)$}{
                    \eIf(\tcp*[f]{collision}){$C^j(t)$}{
                      $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                      \tcp*[f]{randomly switch}
                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB}){
                        $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                      }
                    }{
                      $A^j(t+1) = A^j(t)$
                      \tcp*[f]{stays on the same arm}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
          \label{algo:5:RandTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

@@

\TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(1) \sim \cU([K])$ and $C^j(1)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
              \For{$t = 0, \dots, T-1$}{
                   \uIf(\tcp*[f]{transition $(3)$ or $(5)$}){
                      $A^j(t) \notin \TopM(t)$}
                    {
                      $A^j(t+1) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                      \tcp*[f]{not empty} \\
                      % \tcp*[f]{randomly switch on an arm that had smaller UCB}
                      $s^j(t+1) = \mathrm{False}$
                      \tcp*[f]{aim at an arm with a smaller UCB at $t-1$}
                    }
                    \uElseIf(\tcp*[f]{collision and not fixed}){
                        $C^j(t)$ \emph{and} $\overline{s^j(t)}$}
                      {
                        $A^j(t+1) \sim \cU \left(\TopM(t)\right)$
                        \tcp*[f]{transition $(2)$} \\
                        $s^j(t+1) = \mathrm{False}$
                    }
                    \Else(\tcp*[f]{transition $(1)$ or $(4)$}){
                      $A^j(t+1) = A^j(t)$ \tcp*[f]{stay on the previous arm} \\
                      $s^j(t+1) = \mathrm{True}$ \tcp*[f]{become or stay fixed on a ``chair''}
                    }
                  Play arm $A^j(t+1)$, get new observations (sensing and collision), \\
                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
          \label{algo:5:MCTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

More precisely, under \MCTopM,
% \footnote{This choice is similar to the elementary step used in the \MusicalChair{} algorithm introduced by \cite{Rosenski16}, that gave its name to \MCTopM, even if they restrict to using the empirical averages as indices (\ie, the $0$-greedy algorithm).}
if player $j$ did not encounter a collision when using arm $k$ at time $t$,
then she marks her current arm as a ``chair'' ($s^j(t+1)=\mathrm{True}$),
and will keep using it even if collisions happen in the future (Lines~$9$-$11$).
%
As soon as this ``chair'' $k$ is no longer in $\widehat{M_j}(t)$,
a new arm is sampled uniformly from a subset of $\TopM(t)$,
defined with the previous indices $U^j(t-1)$ (Lines~$3$-$5$).
%
The subset enforces a certain inequality on indices,
$U_{k'}^j(t-1) \leq U^j_{k}(t-1)$ and $U_{k'}^j(t) \geq U^j_{k}(t)$,
when switching from $k=A^j(t)$ to $k'=A^j(t+1)$.
This helps to control the number of such changes of arm,
as shown in Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}.
% the Appendix~\ref{proof:5:collisionsMCTopM}
The considered subset is never empty as it contains
at least the arm replacing the $k\in\TopM(t-1)$ in $\TopM(t)$.
Collisions are dealt with only for non-fixed player $j$,
and when the previous arm is still in $\TopM(t)$.
%
In this case, a new arm is sampled uniformly from $\TopM(t)$ (Lines~$6$-$8$).
%
%
This stationary aspect helps to minimize the number of collisions,
as well as the number of switches of arm.
%
The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM, and they are illustrated in Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below.

\begin{figure}[h!]
  \resizebox{1.05\textwidth}{!}{
  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.0]
      %
      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
      %
      \draw [black,->] (start) -> (notfixed.20);
      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t)$,\\$A^j(t) \in \TopM(t)$} (1);
      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
      %
  \end{tikzpicture}
  }
  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
  \label{fig:5:StateMachineAlgorithm_MCTopM}
\end{figure}

>>>>

% \TODOL{Est-ce qu'on enlève toute mention de la lower-bound et de "order-optimal" du coup ?!}

@@

% \TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(0) \sim \cU([K])$ and $C^j(0)=\mathrm{False}$ \\
              \For{$t = 1, \dots, T$}{
                  %
                  \eIf{$A^j(t-1) \notin \TopM(t)$}{
                    \eIf(\tcp*[f]{collision at previous step}){$C^j(t-1)$}{
                      $A^j(t) \sim \cU \left(\TopM(t)\right)$
                      \tcp*[f]{randomly switch}
                      }(\tcp*[f]{randomly switch on an arm that had smaller UCB}){
                        $A^j(t) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                      }
                    }{
                      $A^j(t) = A^j(t-1)$
                      \tcp*[f]{stays on the same arm}
                    }
                  Play arm $A^j(t)$, get new observations (sensing and collision), \\
                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \RandTopM{} decentralized learning policy]{The \RandTopM{} decentralized learning policy (for an index policy $U^j$).}
          \label{algo:5:RandTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

@@

% \TODOL{Est-ce qu'on devrait utiliser $t=1,\dots,T$ et pas $t=0,\dots,T-1$ et utiliser $A(t)$ et pas $A(t+1)$ pour être consistant avec les autres chapitres ?}

% \vspace*{-5pt}  % XXX remove if problem
% \begin{small}  % XXX remove if problem
  \begin{figure}[h!]
      \begin{framed}  % XXX remove if problem
      % \begin{small}  % XXX remove if problem
      \centering
      % Documentation at http://mirror.ctan.org/tex-archive/macros/latex/contrib/algorithm2e/doc/algorithm2e.pdf if needed
      % Or https://en.wikibooks.org/wiki/LaTeX/Algorithms#Typesetting_using_the_algorithm2e_package
      % \removelatexerror% Nullify \@latex@error % Cf. http://tex.stackexchange.com/a/82272/
      \begin{algorithm}[H]
          % XXX Options
          % \LinesNumbered  % XXX Option to number the line
          % \RestyleAlgo{boxed}
          % XXX Input, data and output
          % \KwIn{$K$ and policy $P^j$ for arms set $[K]$\;}
          % \KwData{Data}
          % \KwResult{Result}
          % XXX Algorithm
              Let $A^j(0) \sim \cU([K])$ and $C^j(0)=\mathrm{False}$ and $s^j(1)=\mathrm{False}$ \\
              \For{$t = 1, \dots, T$}{
                   \uIf(\tcp*[f]{transition $(3)$ or $(5)$}){
                      $A^j(t-1) \notin \TopM(t)$}
                    {
                      $A^j(t) \sim \cU \left(\TopM(t) \cap \left\{k : U_k^j(t-1) \leq U^j_{A^j(t)}(t-1)\right\}\right)$
                      \tcp*[f]{not empty} \\
                      % \tcp*[f]{randomly switch on an arm that had smaller UCB}
                      $s^j(t) = \mathrm{False}$
                      \tcp*[f]{aim at an arm with a smaller UCB at $t-1$}
                    }
                    \uElseIf(\tcp*[f]{collision and not fixed}){
                        $C^j(t-1)$ \emph{and} $\overline{s^j(t-1)}$}
                      {
                        $A^j(t) \sim \cU \left(\TopM(t)\right)$
                        \tcp*[f]{transition $(2)$} \\
                        $s^j(t) = \mathrm{False}$
                    }
                    \Else(\tcp*[f]{transition $(1)$ or $(4)$}){
                      $A^j(t) = A^j(t-1)$ \tcp*[f]{stay on the previous arm} \\
                      $s^j(t) = \mathrm{True}$ \tcp*[f]{become or stay fixed on a ``chair''}
                    }
                  Play arm $A^j(t)$, get new observations (sensing and collision), \\
                  Compute the indices $U^j_k(t+1)$ and set $\TopM(t+1)$ for next step.
              }
              \caption[The \MCTopM{} decentralized learning policy]{The \MCTopM{} decentralized learning policy (for an index policy $U^j$).}
          \label{algo:5:MCTopM}
      \end{algorithm}
      % \end{small}  % XXX remove if problem
      \end{framed}  % XXX remove if problem
  \end{figure}
% \end{small}  % XXX remove if problem
% \vspace*{-5pt}  % XXX remove if problem

More precisely, under \MCTopM,
% \footnote{This choice is similar to the elementary step used in the \MusicalChair{} algorithm introduced by \cite{Rosenski16}, that gave its name to \MCTopM, even if they restrict to using the empirical averages as indices (\ie, the $0$-greedy algorithm).}
if player $j$ did not encounter a collision when using arm $k$ at time $t$,
then she marks her current arm as a ``chair'' ($s^j(t)=\mathrm{True}$),
and will keep using it even if collisions happen in the future (Lines~$9$-$11$).
%
As soon as this ``chair'' $k$ is no longer in $\widehat{M_j}(t)$,
a new arm is sampled uniformly from a subset of $\TopM(t)$,
defined with the previous indices $U^j(t)$ (Lines~$3$-$5$).
%
The subset enforces a certain inequality on indices,
$U_{k'}^j(t-1) \leq U^j_{k}(t-1)$ and $U_{k'}^j(t) \geq U^j_{k}(t)$,
when switching from $k=A^j(t-1)$ to $k'=A^j(t)$.
This helps to control the number of such changes of arm,
as shown in Lemma~\ref{lem:5:elementaryLemma_RandTopM_MCTopM}.
% the Appendix~\ref{proof:5:collisionsMCTopM}
The considered subset is never empty as it contains
at least the arm replacing the $k\in\TopM(t-1)$ in $\TopM(t)$.
Collisions are dealt with only for non-fixed player $j$,
and when the previous arm is still in $\TopM(t)$.
%
In this case, a new arm is sampled uniformly from $\TopM(t)$ (Lines~$6$-$8$).
%
%
This stationary aspect helps to minimize the number of collisions,
as well as the number of switches of arm.
%
The five different transitions $(1)$, $(2)$, $(3)$, $(4)$, $(5)$ refer to the notations used in the analysis of \MCTopM, and they are illustrated in Figure~\ref{fig:5:StateMachineAlgorithm_MCTopM} below.

\begin{figure}[h!]
  \resizebox{1.03\textwidth}{!}{
  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.0]
      %
      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
      \node (notfixed) at (1,0) [draw,rectangle,very thick] {Not fixed, $\overline{s^j(t-1)}$};
      \node (fixed) at (0,0) [draw,rectangle,very thick] {Fixed, $s^j(t-1)$};
      %
      \draw [black,->] (start) -> (notfixed.20);
      \draw [color=cyan,very thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t-1)}, A^j(t-1) \in \TopM(t)$} (fixed);
      \path [color=blue,very thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t-1)$,\\$A^j(t-1) \in \TopM(t)$} (1);
      \path [color=red,very thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t-1) \notin \TopM(t)$} (1);
      \path [color=darkgreen,very thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t-1) \in \TopM(t)$} (fixed);
      \draw [color=red,very thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t-1) \notin \TopM(t)$} (notfixed);
      %
  \end{tikzpicture}
  }
  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
  \label{fig:5:StateMachineAlgorithm_MCTopM}
\end{figure}
